---
title: "Penguin Bill Dimensions by Species"
author: "Nitpreet Bamra"
date: "2026-09-22"
---

## Data Disclosure

The data from this analysis is from the Palmer Penguins dataset. This dataset can be found at the following link:

https://allisonhorst.github.io/palmerpenguins/

This dataset is by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network, and gives information about penguins.

The goal of this analysis is to compare the bill dimensions of penguins across the different penguin species.

```{r}
#| message: false
#| warning: false

library(palmerpenguins)
library(tidyverse)

penguins
```

## Data Exploration

unique species:

```{r}
#| warning: false

unique_species <- unique(penguins['species'])

unique_species
```

unique body mass:

```{r}
#| warning: false

body_mass <- unique(penguins['body_mass_g'])


body_mass
```


## Data Cleaning

first lets remove all rows with either a null species or body_mass_g

```{r}
penguins_clean <- penguins |>
  drop_na(species, body_mass_g)

penguins_clean
```

## Average body mass

```{r}
mass_per_species <- penguins_clean |>
  group_by(species) |>
  summarize(
    mean_body_mass = mean(body_mass_g)
  )

mass_per_species
```


## Bar Graph Visualization 

AI disclosure: The R code used to create the bar plot was generated with assistance from OpenAI’s ChatGPT (accessed September 22, 2026).
OpenAI. (2026). ChatGPT [Large language model]. https://chatgpt.com/

```{r}
ggplot(mass_per_species, aes(x = species, y = mean_body_mass)) +
  geom_col() +
  labs(
    title = "Average Body Mass by Penguin Species",
    x = "Penguin Species",
    y = "Average Body Mass (g)"
  ) +
  theme_minimal()
```
